From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam

### 2025-03-05 manuscript.qmd
Final cleanups

---
title: 'From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam'
authors:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: 'https://haziqj.ml'
    corresponding: true
    affiliations:
      - name: Universiti Brunei Darussalam
        id: ubd
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410
  - name: Amira Barizah Noorosmawie
    affiliations:
      - ref: ubd        
  - name: Hafeezul Waezz Rabu
    affiliations:
      - ref: ubd
  - name: Lutfi Abdul Razak
    orcid: 0000-0003-1681-0895
    email: lutfi.razak@ubd.edu.bn
    affiliations:
      - name: UBD School of Business and Economics
        id: ubdsbe
        # department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410         
date-modified: last-modified
abstract: |
  This article introduces the first publicly available data set for analysing the Brunei housing market, covering more than 30,000 property listings from 1993 to early 2025. The data set, curated from property advertisements in newspapers and online platforms, includes key attributes such as price, location, property type, and physical characteristics, enriched with area-level spatial information. Comprehensive and historical, it complements the Brunei Darussalam Central Bank’s Residential Property Price Index (RPPI), addressing the limitations of restricted access to raw RPPI data and its relatively short timeline since its inception in 2015. Data collection involved manual transcription from archival sources and automated web scraping using programmatic methods, supported by innovative processing with Large Language Models (LLMs) to codify unstructured text. The data set enables spatial and temporal analysis, with potential applications in economics, urban planning, and real estate research. Although listing prices serve only as a proxy for market values and may deviate from actual sale prices due to negotiation dynamics and other factors, this data set still provides a valuable resource for reproducible analysis of housing market trends and for informing policy decisions.
keywords:
  - Housing Market
  - Brunei
  - Property Listings
  - Spatial Data
  - Web Scraping
  - Large Language Models
---

@@

- **First of its kind for Brunei**. This data set is the first publicly available compilation of Brunei’s residential property data, covering `{r} scales::comma(nrow(hsp))` listings from `r min(years)` to early `r max(years)`. Previous studies on Brunei's housing have either relied on recent or non-spatial data [@ng2022characterising;@jamil2025leveraging], or have primarily employed qualitative methods due to the absence of structured data [@hassan2011cultural;@hassan2023sociocultural]. Researchers may use this data to explore temporal variations in property listings and to compare housing characteristics across different regions within Brunei.

- **Historical and spatial insights**. Spanning over three decades, this data set allows for the study of long-term housing trends in Brunei. The spatial information available enables analysis at various administrative levels, providing insights into urban development patterns, regional differences, and the evolution of housing characteristics over time. This historical depth is particularly valuable given the lack of data before the establishment of the Residential Property Price Index (RPPI) [@bdcb2021technical] in 2015.

- **Methodological innovation in data curation**. The data collection process employs a unique combination of manual transcription from archival sources, automated web scraping, and AI-based data cleaning using Large Language Models (LLMs). This open source and multi-method approach enhances reproducibility and transparency in data curation. The present work provides a reproducible framework for assembling complex data sets, which can be adapted for similar data collection efforts in other other domains, thereby advancing best practices in the field.

@@

The housing market is a key indicator of economic health and social well-being, yet comprehensive and publicly accessible data sets in Brunei remain limited. 
As far as we are aware, this is the first data set of its kind in Brunei, motivated by the need to fill the gap in publicly available data on the local housing market.

@@

Brunei’s private residential property market offers a variety of options, including detached houses, townhouses, and apartments [@hassan2023sociocultural]. 
Based on this, property types have been categorised into four main groups–Detached, Semi-Detached, Terrace, and Apartment–to accurately reflect the diversity of property types in Brunei.
Additionally, there are a small number of records that reflect listings for land, which are categorised accordingly as "Land".

@@

The date variable in the data set refers to the date on which the property listing was obtained. 
It is important to note that this is not the date the property was sold, nor does it necessarily reflect the precise timing of any transactions related to the property. 
Users should set their expectations accordingly, as the primary purpose of the date is to capture the state of the housing market at a specific point in time.

For analysis, we recommend aggregating data by quarters, as represented by the `quarter` variable. 
This aggregation helps address potential issues like missing data (see subsection below) and provides a more stable and robust representation of market trends, making it suitable for temporal analysis of the housing market.

@@

In any data collection effort, it is not surprising to encounter missing data. 
Likewise for our data set, we faced the challenge of missing values across several property characteristics--such as plot area, floor area, beds, baths, and others.
A preliminary analysis of the missing data patterns indicates that the missingness is not completely random, with certain variables displaying dependencies on others.
Advertisers often include only the information they deem most marketable or necessary, while other details may be omitted if they are considered standard or implied. 
For instance, a listing might specify the square footage and price but leave out the number of bedrooms and bathrooms, assuming that prospective buyers can infer these details.

@@

1. **Outlier detection.** Summary statistics analyses to identify and flag anomalous values. For instance, a built-up area recorded as 0.1 square feet or beds and baths exceeding 20 is suspect, and highly likely points towards human error or model hallucinations. Such anomalies were flagged and manually reviewed for correction.
2. **Internal consistency checks.** This made use of substantive knowledge about Brunei's housing market [@jamil2025spatiotemporal]. An example is using the price per square foot indicator, whose value typically falls within a known range. Therefore any deviations were scrutinised to ensure that variables such as price and floor area were correctly recorded. This was similarly applied to other variables such as plot area, beds, and baths.
3.	**Duplicate records detection.** We also performed duplicate record detection to identify any repeated entries that might have arisen from overlapping data sources or transcription errors. Any duplicates that were identified were carefully reviewed and removed to ensure that each property listing was uniquely represented in the span of one calendar month.

@@

OpenAI's models topped the accuracy charts, with the o1-mini and gpt-40 models achieving the accuracy scores of 99.2% and 98.9% respectively.
The reasoning model deepseek-r1:14b was the best performing locally run model, scoring 96.9% accuracy.
Evidently, the smaller the model, the less accurate the extraction process (cf. llama3.2 70.8%).
Most inaccuracies were seen in the `status` variable, where the models had a difficult time parsing the correct build status from the vague verbose advertisement.
However, key variables such as `price`, `type`, `plot_area`, `floor_area`, `beds`, and `baths` are generally reliable (>90% accuracy).

>>>>

---
title: 'From Archives to AI: Residential Property Data Across Three Decades in Brunei Darussalam'
authors:
  - name: Haziq Jamil
    orcid: 0000-0003-3298-1010
    email: haziq.jamil@ubd.edu.bn
    url: 'https://haziqj.ml'
    corresponding: true
    affiliations:
      - name: Universiti Brunei Darussalam
        id: ubd
        department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410
  - name: Amira Barizah Noorosmawie
    affiliations:
      - ref: ubd        
  - name: Hafeezul Waezz Rabu
    affiliations:
      - ref: ubd
  - name: Lutfi Abdul Razak
    orcid: 0000-0003-1681-0895
    email: lutfi.razak@ubd.edu.bn
    affiliations:
      - name: UBD School of Business and Economics
        id: ubdsbe
        # department: Mathematical Sciences, Faculty of Science
        address: Universiti Brunei Darussalam, Jalan Tungku Link
        city: Bandar Seri Begawan
        country: Brunei
        postal-code: BE 1410         
date-modified: last-modified
abstract: |
  This article introduces the first publicly available data set for analysing the Brunei housing market, covering more than 30,000 property listings from 1993 to early 2025. The data set, curated from property advertisements in newspapers and online platforms, includes key attributes such as price, location, property type, and physical characteristics, enriched with area-level spatial information. Comprehensive and historical, it complements the Brunei Darussalam Central Bank’s Residential Property Price Index (RPPI), addressing the limitations of restricted access to raw RPPI data and its relatively short timeline since its inception in 2015. Data collection involved manual transcription from archival sources and automated web scraping using programmatic methods, supported by innovative processing with Large Language Models (LLMs) to codify unstructured text. The data set enables spatial and temporal analysis, with potential applications in economics, urban planning, and real estate research. Although listing prices are only a proxy for market values and may deviate from actual sale prices due to negotiation dynamics and other factors, this data set still provides a valuable resource for quantitative analyses of housing market trends and for informing policy decisions.
keywords:
  - Housing Market
  - Brunei
  - Property Listings
  - Spatial Data
  - Web Scraping
  - Large Language Models
---

@@

- **A first-of-its-kind data set for Brunei**. This data set is the first publicly available compilation of Brunei’s residential property data, covering `{r} scales::comma(nrow(hsp))` listings from `r min(years)` to early `r max(years)`. Previous studies on Brunei's housing have either relied on recent or non-spatial data [@ng2022characterising;@jamil2025leveraging], or have primarily employed qualitative methods due to the absence of structured data [@hassan2011cultural;@hassan2023sociocultural]. Researchers may use this data to explore temporal variations in property listings and to compare housing characteristics across different regions within Brunei.

- **Historical and spatial insights**. Spanning over three decades, this data set enables analysis of long-term housing trends at various administrative levels in Brunei, providing insights into urban development patterns, regional differences, and the evolution of housing characteristics over time. This historical depth is particularly valuable given the lack of data prior to the establishment of the Residential Property Price Index (RPPI) [@bdcb2021technical] in 2015.

- **Methodological innovation in data curation**. The data collection process employs a unique combination of manual transcription from archival sources, automated web scraping, and AI-based data cleaning using Large Language Models (LLMs). This multi-method approach offers a reproducible framework for assembling complex data sets, which can be adapted for similar data collection efforts in other other domains, thereby advancing best practices in the field.

@@

The housing market is a key indicator of economic health and social well-being, yet comprehensive and publicly accessible data sets in Brunei remain limited. 
To the best of our knowledge, this is the first data set of its kind in Brunei, motivated by the need to fill the gap in publicly available local housing market data.

@@

Brunei’s private residential property market offers a variety of options, including detached houses, townhouses, and apartments [@hassan2023sociocultural]. 
Based on this diversity, property types have been categorised into four main groups--Detached, Semi-Detached, Terrace, and Apartment--to accurately reflect the range of residential properties in Brunei.
Additionally, there are a small number of records that reflect listings for land, which are categorised accordingly as "Land".

@@

The `date` variable represents the date on which the property listing was obtained, rather than the date of sale or transaction, and should be interpreted as a snapshot of market conditions at a given point in time.
For analysis, we recommend aggregating data by quarters, as represented by the `quarter` variable. 
This aggregation helps address potential issues like missing data (see subsection below) and provides a more stable and robust representation of market trends, making it suitable for temporal analysis of the housing market.

@@

In any data collection effort, it is unsurprising to encounter missing data. 
Likewise in our data set, missing values occur across several property characteristics--such as plot area, floor area, beds, baths, and others.
A preliminary analysis of the missing data patterns indicates that the missingness is not completely random, with certain variables displaying dependencies on others.
Advertisers often include only the information they deem most marketable or necessary, while other details may be omitted if they are considered standard or implied. 
For instance, a listing might specify the square footage and price but leave out the number of bedrooms and bathrooms, assuming that prospective buyers are able to infer these details.

@@

1. **Outlier detection.** Summary statistics analyses to identify and flag anomalous values. For instance, a built-up area recorded as 0.1 square feet or an implausibly high number of beds and baths (e.g., >20) would highly likely indicate a possible error, prompting manual review. Such anomalies were flagged and manually reviewed for correction.
2. **Internal consistency checks.** This made use of substantive knowledge about Brunei's housing market [@jamil2025spatiotemporal]. An example is using the price per square foot indicator, whose value typically falls within a known range. Therefore any deviations were scrutinised to ensure that variables such as price and floor area were correctly recorded. This was similarly applied to other variables such as plot area, beds, and baths.
3.	**Duplicate records detection.** We also performed duplicate record detection to identify any repeated entries that might have arisen from overlapping data sources or transcription errors. Any duplicates that were identified were carefully reviewed and removed to ensure that each property listing was uniquely represented in the span of one calendar month.

@@

OpenAI's models topped the accuracy charts, with the o1-mini and gpt-40 models achieving the accuracy scores of 99.2% and 98.9% respectively.
The reasoning model deepseek-r1:14b was the best performing locally run model, scoring 96.9% accuracy.
Evidently, the smaller the model, the less accurate the extraction process (cf. llama3.2 70.8%).
Most inaccuracies occurred in the `status` variable, where models struggled to parse the correct build status from vague advertisement descriptions.
However, key variables such as `price`, `type`, `plot_area`, `floor_area`, `beds`, and `baths` are generally reliable (>90% accuracy).
